Abstract

### 2025-05-22 output/Letter_to_editor_reviewers.Rmd
Add example of orthogonal (miss)-interpretation

There are several references in the literature that have found contradictory results that, we believe, are due to disregarding the latent covariance structure (i.e., using Reckase & McKinley's formulae in a non-orthonormal space).
For example, 
<!-- # TODO: Example(s) from Gecko search -->.
We believe this is also what happens when MIRT software such as NOHARM tries to force a correlated latent trait structure into its default orthogonal solution, as you have noted yourself [@ackerman_multidimensional_2005-1, p. 7].
<!-- # TODO: Correctly expressed? -->
While we consider this to be out of the scope of the current paper, we will gladly add further details to explain why these results are contradictory and how our method may solve the problem.

>>>>

There are several references in the literature that have found contradictory results that, we believe, are due to disregarding the latent covariance structure (i.e., using Reckase & McKinley's formulae in a non-orthonormal space).
For example, @min_evaluation_2007 [p. 49] computes the item parameters in the M2PL from a priori fixed values of $MDISC_i$ and $MDIFF_i$ (the name they use for $MID_i$), using @reckase_difficulty_1985 and @reckase_discriminating_1991's formulae.
They find that "[...] bias [...] was more apparent when ability distribution was separated from the
default distribution, such as was the case in groups 2 to 5".
Although they attribute this to a possible "degree of confounding among errors of parameter estimation and scale
transformation", their groups 2 to 5 have all correlations equal (or almost equal) to .5, as it happens.

We believe this is also what happens when MIRT software such as NOHARM tries to force a correlated latent trait structure into its default orthogonal solution, as this editor himself has noted before [@ackerman_multidimensional_2005-1, p. 11].
While we consider this to be out of the scope of the current paper, we will gladly add further details to explain why these results are contradictory and how our method may solve the problem.
